---
title: "01 Data Analysis"
output:
  html_document:
    df_print: paged
---

The following exercises are part of a course/workshop to learn basic R with R Studio and Git ([Syllabus](https://docs.google.com/document/d/1O_v-VTRHChedoDvVtzn8xNG0jBqJxMW0Q7-0bJ90Q48/edit))


## Setup

Data imported and cleaned with R script [data_import.R](./Data/data_import.R)

```{r setup}

# by making an object with a list of packages, we can later call the citations of these packages (see section 'References' below)

packages <- c('tidyverse', 
              'pacman')


# here we can load the packages. p_load from the package pacman checks if the package is already downloaded and if so loads it, if not downloads

pacman::p_load(packages, character.only = T)


# here we load the data

df <- readRDS("C:/Github/reproducible_R_Git/Data/df.RDS")
```

## Linear Regressions

### Store models

We run regressions of life expectancy on GDP, region and variables that categorize the mottos of different countries. We store these regressions in objects named model1, model2, etc.

```{r ols}

# metric (original measures)

model1 <- lm(life_exp ~ gdppc_k + region, data = df)
model2 <- lm(life_exp ~ gdppc_k + has_god + has_unity, data = df)


# XY-standardized
## For dichotomous variables, it is not necessary to standardize X (depends on what interpretation you are looking for)

## standardized

df$std_life_exp <- as.numeric(scale(df$life_exp, center = TRUE, scale = TRUE))
df$std_gdppc <- as.numeric(scale(df$gdppc_k, center = TRUE, scale = TRUE))

## standardized models
model1B <- lm(std_life_exp ~ std_gdppc + region, data = df)
model2B <- lm(std_life_exp ~ std_gdppc + has_god + has_unity, data = df)
```

### Tabulate

#### Create

Create a table of results using base R

```{r ols1}
sum_tab <- list(summary(model1)$coefficients,
                summary(model1B)$coefficients,
                summary(model2)$coefficients,
                summary(model2B)$coefficients)

# name the items in the list
names(sum_tab) <- c("model1", "model1B", "model2", "model2B")

# calculate p-values
## create tables first as dataframes
tbl1 <- as.data.frame(sum_tab$model1)
tbl2 <- as.data.frame(sum_tab$model1B)
tbl3 <- as.data.frame(sum_tab$model2)
tbl4 <- as.data.frame(sum_tab$model2B)

## remove SE and t-values, and make one table

tbl_all <- as.data.frame(matrix(nrow = 25, ncol = 9))

colnames(tbl_all) <- c("Variable", "m1_b", "m1_p", "m1B_b", "m1B_p", "m2_b", "m2_p", "m2B_b", "m2B_p")

tbl_all$Variable[1:23] <- rownames(tbl1)
tbl_all$Variable[24:25] <- c("Motto God", "Motto Unity")
tbl_all$m1_b[1:23] <- round(tbl1$Estimate,2)
tbl_all$m1B_b[1:23] <- round(tbl2$Estimate,2)
tbl_all$m2_b[c(1,2,24,25)] <- round(tbl3$Estimate,2)
tbl_all$m2B_b[c(1,2,24,25)] <- round(tbl4$Estimate,2)

tbl_all$m1_p[1:23] <- round(tbl1$`Pr(>|t|)`,2)
tbl_all$m1B_p[1:23] <- round(tbl2$`Pr(>|t|)`,2)
tbl_all$m2_p[c(1,2,24,25)] <- round(tbl3$`Pr(>|t|)`,2)
tbl_all$m2B_p[c(1,2,24,25)] <- round(tbl4$`Pr(>|t|)`,2)

tbl_all[is.na(tbl_all)] <- ""

# remove the word "region" from all labels
tbl_all <- data.frame(lapply(tbl_all, function(x) {
  gsub("region","",x)
}))
```

#### Print

```{r printdf}

print.data.frame(tbl_all, row.names = F)

```

