---
title: "01 Data Analysis"
output: html_notebook
---

The following exercises are part of a course/workshop to learn basic R with R Studio and Git ([Syllabus](https://docs.google.com/document/d/1O_v-VTRHChedoDvVtzn8xNG0jBqJxMW0Q7-0bJ90Q48/edit))


## Setup

Data imported and cleaned with R script [data_import.R](./Data/data_import.R)

```{r setup}

# by making an object with a list of packages, we can later call the citations of these packages (see section 'References' below)

packages <- c('tidyverse', 
              'pacman')


# here we can load the packages. p_load from the package pacman checks if the package is already downloaded and if so loads it, if not downloads

pacman::p_load(packages, character.only = T)


# here we load the data

df <- readRDS("C:/Github/reproducible_R_Git/Data/df.RDS")
```

## Linear Regressions

### Store models

We run regressions of life expectancy on GDP, region and variables that categorize the mottos of different countries. We store these regressions in objects named model1, model2, etc.

```{r ols}
model1 <- lm(life_exp ~ gdppc_k + region, data = df)
model2 <- lm(life_exp ~ gdppc_k + has_god + has_unity, data = df)
```

### Tabulate

```{r ols1}

summary(model2)
```

