---
title: "01 Data Analysis"
output:
  html_document:
    df_print: paged
---

The following exercises are part of a course/workshop to learn basic R with R Studio and Git ([Syllabus](https://docs.google.com/document/d/1O_v-VTRHChedoDvVtzn8xNG0jBqJxMW0Q7-0bJ90Q48/edit))


## Setup

Data imported and cleaned with R script [data_import.R](./Data/data_import.R)

```{r setup}

# by making an object with a list of packages, we can later call the citations of these packages (see section 'References' below)

packages <- c('tidyverse', 
              'pacman')


# here we can load the packages. p_load from the package pacman checks if the package is already downloaded and if so loads it, if not downloads

pacman::p_load(packages, character.only = T)


# here we load the data

df <- readRDS("C:/Github/reproducible_R_Git/Data/df.RDS")
```

## Recoding

As an exercise, lets learn to do some recoding.

For starters, recode life expectancy into ordinal categories. We can do this as numbers or as strings (characters) using `mutate` and `case_when`.

### Ordered Categories

```{r recodes}

df <- df %>%
  mutate(
    life_exp_cat = case_when( # first make numerical recodes
      life_exp <= 60 ~ 1,
      life_exp > 60 & life_exp <= 70 ~ 2,
      life_exp > 70 & life_exp <= 80 ~ 3,
      life_exp > 80 ~ 4,
      .default = NA
      ),
    life_exp_cat_chr = case_when( # now make character recodes
      life_exp <= 60 ~ "60 or less",
      life_exp > 60 & life_exp <= 70 ~ "60 to 70",
      life_exp > 70 & life_exp <= 80 ~ "70 to 80",
      life_exp > 80 ~ "80 or more",
      .default = NA)
    )
```

### String to numeric

Use `case_match` to divide the character based variable back into a numeric variable. 

```{r case_match}

df <- df %>%
  mutate(
    life_exp_cat_num = case_match(
      life_exp_cat_chr,
      "60 or less" ~ 1,
      "60 to 70" ~ 2,
      "70 to 80" ~ 3,
      "80 or more" ~ 4,
      .default = NA
    )
  )

# test that life_exp_cat_num and life_exp_cat are idenitcal
cor(df$life_exp_cat, df$life_exp_cat_num, use = "pairwise.complete.obs")

```

## Linear Regressions

### Store models

We run regressions of life expectancy on GDP, region and variables that categorize the mottos of different countries. We store these regressions in objects named model1, model2, etc.

```{r ols}

# metric (original measures)

model1 <- lm(life_exp ~ gdppc_k + region, data = df)
model2 <- lm(life_exp ~ gdppc_k + has_god + has_unity, data = df)
model3 <- lm(life_exp ~ gdppc_k + region + has_god + has_unity, data = df)

# XY-standardized
## For dichotomous variables, it is not necessary to standardize X (depends on what interpretation you are looking for)

## standardized

df$std_life_exp <- as.numeric(scale(df$life_exp, center = TRUE, scale = TRUE))
df$std_gdppc <- as.numeric(scale(df$gdppc_k, center = TRUE, scale = TRUE))

## standardized models
model1B <- lm(std_life_exp ~ std_gdppc + region, data = df)
model2B <- lm(std_life_exp ~ std_gdppc + has_god + has_unity, data = df)
model3B <- lm(std_life_exp ~ std_gdppc + region + has_god + has_unity, data = df)
```

### Tabulate

#### Create

Create a table of results using base R

```{r ols1}
sum_tab <- list(summary(model1)$coefficients,
                summary(model1B)$coefficients,
                summary(model2)$coefficients,
                summary(model2B)$coefficients,
                summary(model3)$coefficients,
                summary(model3B)$coefficients)

# name the items in the list
names(sum_tab) <- c("model1", "model1B", "model2", "model2B", "model3", "model3B")

# calculate p-values
## create tables first as dataframes
tbl1 <- as.data.frame(sum_tab$model1)
tbl2 <- as.data.frame(sum_tab$model1B)
tbl3 <- as.data.frame(sum_tab$model2)
tbl4 <- as.data.frame(sum_tab$model2B)
tbl5 <- as.data.frame(sum_tab$model3)
tbl6 <- as.data.frame(sum_tab$model3B)

## remove SE and t-values, and make one table

tbl_all <- as.data.frame(matrix(nrow = 25, ncol = 10))

colnames(tbl_all) <- c("Variable", "m1_b", "m1_p", "m1_B", "m2_b", "m2_p", "m2_B", "m3_b", "m3_p", "m3_B")

tbl_all$Variable[1:25] <- rownames(tbl5)
tbl_all$m1_b[1:23] <- round(tbl1$Estimate,2)
tbl_all$m1_B[1:23] <- round(tbl2$Estimate,2)
tbl_all$m2_b[c(1,2,24,25)] <- round(tbl3$Estimate,2)
tbl_all$m2_B[c(1,2,24,25)] <- round(tbl4$Estimate,2)
tbl_all$m3_b <- round(tbl5$Estimate,2)
tbl_all$m3_B <- round(tbl6$Estimate,2)

tbl_all$m1_p[1:23] <- round(tbl1$`Pr(>|t|)`,2)
tbl_all$m2_p[c(1,2,24,25)] <- round(tbl3$`Pr(>|t|)`,2)
tbl_all$m3_p[] <- round(tbl5$`Pr(>|t|)`,2)

tbl_all[is.na(tbl_all)] <- ""


# remove the word "region" from all labels
tbl_all <- data.frame(lapply(tbl_all, function(x) {
  gsub("region","",x)
}))
```

#### Print

```{r printdf}

print.data.frame(tbl_all, row.names = F)

```

## Citations

```{r cite}
packages %>%
  map(citation) %>%
  print(style = "text")
```
